A Vectorized Correlation Function Estimator For Large-Scale Structure

### 2018-11-15 paper/paper.tex
small abstract edits

% aastex parameters
%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
\received{not yet but hopefully soon!!}
%\revised{not yet}
%\accepted{not yet}
%\submitjournal{ApJ}
\shorttitle{tensor estimator for large-scale structure}
\shortauthors{storey-fisher and hogg}

@@

\begin{abstract}\noindent
% Context
The two-point correlation function (2PCF) is one of the most commonly used statistics to characterize clustering in galaxy surveys. However, current estimators of the 2PCF have significant limitations. The standard Landy-Szalay estimator evaluates the 2PCF in bins of radial separation between objects, and the choice of bins introduces a trade-off between bias and variance. Further trade-offs are required to determine the dependence of clustering on galaxy properties, typically by subsampling the objects. A marked correlation function, which weights the 2PCF based on some property, avoids some of these issues but still requires binning and contains a bias with sample selection.
% Aims
We present a new, vectorized estimator for the 2PCF that resolves these issues. 
% Methods
It projects pairs of objects onto any set of continuous basis functions, eliminating the need for binning. These functions can also depend on galaxy properties, so the estimator effectively computes a regression against these properties. This produces a vector that encodes the clustering information, and the 2PCF can be quickly evaluated at any value of the given property, giving a precise, unbiased characterization of the evolution.
% Results
We apply this estimator to the main galaxy sample of the Sloan Digital Sky Survey and demonstrate that it precisely characterizes the luminosity evolution of clustering. We also use the estimator to directly fit for the baryon acoustic feature and show that it extracts Fisher-optimal statistics. These results indicate that this new method provides the accuracy and precision needed to analyze upcoming galaxy surveys.

>>>>

% aastex parameters
%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
\received{not yet but hopefully soon!!}
%\revised{not yet}
%\accepted{not yet}
%\submitjournal{ApJ}
\shorttitle{vectorized estimator for large-scale structure}
\shortauthors{storey-fisher and hogg}

@@

\begin{abstract}\noindent
% Context
The two-point correlation function (2PCF) is one of the most commonly used statistics to characterize clustering in galaxy surveys. However, current estimators of the 2PCF have significant limitations. The standard Landy-Szalay estimator evaluates the 2PCF in bins of radial separation between objects, and the choice of bins introduces a trade-off between bias and variance. Further trade-offs are required to determine the dependence of clustering on galaxy properties, typically by subsampling the objects. A marked correlation function, which weights the 2PCF based on some property, avoids some of these issues but still requires binning and contains a bias with sample selection.
% Aims
We present a new, vectorized estimator for the 2PCF that resolves these issues. 
% Methods
It generalizes the Landy-Szalay estimator to depend on a tensor of random catalog pairs. The estimator then projects pairs onto any set of continuous basis functions, eliminating the need for binning. These functions can also depend on galaxy properties, so the estimator effectively computes a regression against these properties. This produces a vector that encodes the clustering information, and the 2PCF can be quickly evaluated at any value of the given property, giving a precise, unbiased, and smooth characterization of the evolution.
% Results
We apply this estimator to the main galaxy sample of the Sloan Digital Sky Survey and demonstrate that it precisely characterizes the luminosity evolution of clustering. We also use the estimator to directly fit for the baryon acoustic feature and show that it extracts Fisher-optimal statistics. These results indicate that this new method provides the accuracy and precision needed to analyze upcoming galaxy surveys.
